---
title: "Iterations and list cols  "
author: "My An Huynh"
date: "2024-10-31"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)

```
### For loops 
basic structure is there's a list for the input and the output will output a list. 

### Loop functions 
are better because you avoid having to define output object. 

### Map functions
Clarify the loop process 
The basic structure is `output = map(input, f)`

Map will take an output and return a list 
If you know what kind of output your function will produce, you can use a specific map variant to help prevent errors and simplify outputs 
   - `map_dbl`
   - `map_lgl`
   - `map_df`
   
If you need to iterate over two inputs, you can use map variants to give two input lists/vectors 
    - `map2`
    - `map2_dbl`
    - `map_df`
    
### Process 
     - Write a single example for fixed inputs
     - Abstract exampe to a function 
     - Embed function in a loop 
     - Re-write using a map statement 
This helps make each step clear, prevent mistakes and only adds complexity when needed

## Here's some lists 

Different objects inside a list. Can access by index or number entry using square brackets, like `l[["1]]`
```{r}
l = list(
  vec_numeric = 1:4,
  unif_sample = runif(1:100),
  mat = matrix(1:8, nrow = 2, ncol = 4, byrow = TRUE),
  summary = summary(rnorm(1000))
)

l[["mat"]][1, 3]
```


Make a list that's hopefully a bit more useful. List of normal samples. Again, access by using `list_norm[["a]]`

```{r}
list_norm = 
  list(
    a = rnorm(20, 0, 5),
    b = rnorm(20, 4, 5),
    c = rnorm(20, 0, 10),
    d = rnorm(20, 4, 10)
  )

list_norm[["a"]]

```

Reuse the function we wrote last time. 

```{r}
mean_and_sd = function(x) {
  
  mean_x = mean(x)
  sd_x = sd(x)
  
  out_df = 
    tibble(
      mean = mean_x,
      sd = sd_x
    )
  
  return(out_df)
}
```

Let's use the function to take mean and sd of all samples

```{r}
mean_and_sd(list_norm[["a"]])
mean_and_sd(list_norm[["b"]])
mean_and_sd(list_norm[["c"]])
mean_and_sd(list_norm[["d"]])

```

## Use a for loop

Create output list, and run a for loop
We have 4 lists of normal (a,b,c,d). For any of those, we have an output list to catch each output separately 

```{r}
output = vector("list", length = 4)

for (i in 1:4)  {
  
  output[[i]] = mean_and_sd(list_norm[[i]])
}

output
```


## Do the same thing 

but with `map` instead 
Map takes input list and then take the function (like mean_and_sd) to run to the list

```{r}
output = map(list_norm, mean_and_sd)

```

Let's do a couple of other things. Map dbl simplifies the output from a list.

Map dfr simplifies the output into a dataframe

```{r}
output = 
  map(list_norm, mean_and_sd) |> 
  bind_rows()

output = map_dfr(list_norm, median)
output = map_dbl(list_norm, IQR)
```

## List columns
Tidy everything up. 
Put my list into this column of sample
```{r}
listcol_df = 
  tibble(
    name = c("a", "b", "c", "d"),
    samp = list_norm
  )

listcol_df

listcol_df |> 
  filter(name %in% c("a", "b"))

listcol_df |> 
  select(-samp)
```

dataframe can also be a list at the same time
```{r}
listcol_df[["samp"]][["a"]]
```

Compute mean and sd 
`listcol_df[["samp"]]` is just extracting a list
```{r eval = FALSE}
mean_and_sd(listcol_df[["samp"]][["a"]])
mean_and_sd(listcol_df[["samp"]][["b"]])

map(listcol_df[["samp"]][["a"]])
```

add a list column. save the column with the output in df. Unnest takes a nested dataframe and unnesting it. 

```{r}
listcol_df |> 
  mutate(
    output = map(samp, mean_and_sd),
    iqr = map_dbl(samp, IQR)) |> 
  select(-samp) |> 
  unnest(output)
    
```

### NSDUH 
Have a df that has the drugs we're looking at in one dataset. 

This is a version of the NSDUH function last time

```{r}
drug_import = function(html, nth) {
  
  out_table = 
    html |> 
    html_table() |> 
    nth(nth) |> 
    slice(-1) |> 
    select(-contains("P Value"))
  
  return(out_table)
}

```

We need to import the html, then extract the correct tables

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)

```

```{r}
drug_import(html = nsduh_html, nth = 1)
drug_import(html = nsduh_html, nth = 4)
drug_import(html = nsduh_html, nth = 5)
```

Column with marj name, table 1/4/5, and the list column that comes with the df 
First, make a tibble with the three columns
```{r}
nsduh_df = 
  tibble(
    drug = c("marj", "cocaine", "heroin"),
    table_n = c(1,4,5),
  ) |> 
  mutate(
    table = map(table_n, drug_import, html = nsduh_html)
    ) |> 
  unnest(table)

nsduh_df |> 
  filter(State == "New York")
```


Create a list column 

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

List column for each location (NY, HI, WA)
```{r}
weather_nest = 
  weather_df |> 
  nest(data = date:tmin)
```

```{r}
weather_nest[["data"]][[1]]
```

Let's try regressing tmax on tmin 

```{r}
lm(tmax ~ tmin, data = weather_nest[["data"]][[1]])
lm(tmax ~ tmin, data = weather_nest[["data"]][[2]])
lm(tmax ~ tmin, data = weather_nest[["data"]][[3]])
```


```{r}
weather_nest |> 
  mutate(model_fit = map(data, \(x) lm(tmax ~ tmin, data = x)))
```

Let's define a function that fits the regression I want. weather_lm is the function name with input of df, and fits lm tmax against tmin. Each time you're running this, the input is a new df (1,2,3)

```{r}
weather_lm = function(df) {
  
  lm(tmax~tmin, data = df)
  
}
```

I could run the model using the short function 

```{r}
weather_lm(weather_nest[["data"]][[1]])
```

```{r}
weather_nest |> 
  mutate(model_fit = map(data, weather_lm))

```

The above is the same as this one:

```{r}
weather_nest |> 
  mutate(model_fit = map(data, \(x) lm(tmax ~tmin, data = x)))
```

